Active Learning for SAT Solver Benchmarking

### 2023-01-23 paper/paper.tex
Add Acknowledgments
Re-phrase evaluation a bit (without deleting content) to make space.

Having selected the most promising hyper-parameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
The aforementioned range $\delta \in \left[0.2, 0.8\right]$ only results in two distinct configurations.
The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking stopping criterion.
It can predict a new solver's PAR-2 ranking with about \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking stopping criterion.
It can predict a new solver's PAR-2 ranking with about \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$), while only needing \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).

@@

Table~\ref{tab:fulldataset} shows how both active-learning approaches compare against static baselines.
The first column shows the two aforementioned AL configurations.
We compare them against a random baseline and a static benchmark set consisting of the instances most frequently sampled by active learning.
The random baseline (\emph{Random}) samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
We repeat this procedure 1000 times and report average results.
The benchmark set of the most frequently AL-sampled instances (\emph{Most Freq.}) uses the average sampling frequencies of instances over all solvers and all Pareto-optimal active-learning approaches.

Our two AL approaches both perform better than random sampling.
However, differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$.
This performance difference between random and non-random sampling also changes for different fractions of sampled runtime (cf. Fig.~\ref{fig:annitraincolorselection}).
A clear advantage of our approach is, though, that it indicates when to stop adding further instances, depending on the trade-off parameter~$\delta$.
While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.

@@

>>>>

Having selected the most promising hyper-parameters, we run our active-learning experiments on the complete Anniversary Track dataset (5355 instances).
The aforementioned range $\delta \in \left[0.2, 0.8\right]$ only results in two distinct configurations.
The best-performing approach for $\delta \in \left[0.2, 0.7\right]$ uses the predicted runtime-label ranking, information-gain sampling, and ranking stopping criterion.
It can predict a new solver's PAR-2 ranking with \SI{90.48}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{5.41}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).
The best-performing approach for $\delta \in (0.7, 0.8]$ uses the predicted runtime-label ranking, uncertainty sampling, and ranking stopping criterion.
It can predict a new solver's PAR-2 ranking with \SI{92.33}{\%} accuracy ($O_{\operatorname{acc}}$) in only \SI{10.35}{\%} of the full evaluation time ($O_{\operatorname{rt}}$).

@@

Table~\ref{tab:fulldataset} shows how both active-learning approaches (column \emph{AL}) compare against two static baselines:
\emph{Random} samples instances until it reaches roughly the same fraction of runtime as the AL benchmark sets.
We repeat sampling 1000 times and report average results.
\emph{Most Freq.} uses the instances most frequently sampled by active learning.
In particular, we consider the average sampling frequencies over all solvers and Pareto-optimal active-learning approaches.

Both our AL approaches perform better than random sampling.
However, the performance differences are not significant regarding a Wilcoxon signed-rank test with $\alpha = 0.05$ and also depend on the fraction of sampled runtime (cf. Fig.~\ref{fig:annitraincolorselection}).
A clear advantage of our approach is, though, that it indicates when to stop adding further instances, depending on the trade-off parameter~$\delta$.
While the active-learning results are less strong on the full dataset than on the smaller tuning dataset, they still show the benefit of making benchmark selection dependent on the solvers to distinguish.

@@

\subsubsection*{Acknowledgments.}

This work was supported by the Ministry of Science, Research and the Arts Baden-Württemberg, project \emph{Algorithm Engineering for the Scalability Challenge (AESC)}.
